### 2024-01-03 paper/main.tex

\documentclass[aps,twocolumn,prl,showpacs,showkeys,preprintnumbers,superscriptaddress,nobibnotes,floatfix,longbibliography,notitlepage,nofootinbib]{revtex4-2}

\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{appendix}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{comment}
\usepackage{bbold}
\usepackage[shortlabels]{enumitem}
\usepackage{color}
\usepackage{slashed}
\usepackage{subfigure}
\usepackage{setspace}
\usepackage{footnote}
\usepackage{lipsum}
\usepackage{multirow}
%\usepackage{longtable}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}
\usepackage[capitalize]{cleveref}
\usepackage{braket}
\usepackage[compat=1.1.0]{tikz-feynman}
\usepackage{multirow}
\usepackage{physics}
\usepackage{feynmp-auto}
\usepackage[normalem]{ulem}
\usepackage{url}
\usepackage{units}
\usepackage[normalem]{ulem}
\newcommand{\yy}[1]{\color{red}[YY:#1]\color{black}}

\input{sections/packages}
\begin{document}

\input{paper/sections/preamble}
\input{paper/sections/introduction}
\input{paper/sections/model}
\input{paper/sections/detector_response}
\input{paper/sections/stats}
\input{paper/`sections/conclusion}

\bibliography{ic_sn_hnl}

\end{document}

% Outline
- Introduction: 
- Detector response
  - What kind of noise, Standard neutrino
  - Atm. background reduction ?
- SNEWPY treatment for general models the pipeline etc.
  - The two models that we have condsidered Lagrangians
  - fluxes and results for them
- Conclusion
  - IC is useful for SN
  - Use our software