### 2024-01-05 paper/sections/conclusion.tex
Updates from Overleaf

\textbf{\textit{Results}}---
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_sensitivity}
    \caption{\textbf{\textit{Exclusion limit for the majoron case.}}
    }
    \label{fig:sensitivity}
\end{figure}
We present the exclusion limit in \fig{sensitivity}.

\textbf{\textit{Outlook}}---